---
title: "Generate Random Sampling Points by Minimum Distance"
author: "Anthony Caravaggi"
date: "`r format(Sys.Date())`"
output: github_document
---

Generate Random Sampling Points by Minimum Distance
================
Anthony Caravaggi
18 April 2018

Libraries

```{r, message = FALSE}
library(ggmap)
library(spatstat)
library(rgdal)
library(rgeos)
library(raster)
```
This demonstration will go through the generation of random points separated by a minimum distance threshold, step-by-step. A function, Rpoints, condensing the process can be found at the bottom of the page. 

Create bounding polygon.

``` r
x <- c(-3.541238, -3.547341, -3.561397, -3.572132, -3.572495, -3.551687, -3.566528, -3.599789, -3.592162,
       -3.604156, -3.629145, -3.618190, -3.585571, -3.572779, -3.525398, -3.504483, -3.525454, -3.541238)
y <- c(51.68106, 51.69050, 51.7003, 51.69617, 51.68800, 51.67405, 51.64949, 51.65832, 51.67560,
       51.68177, 51.71373, 51.72463, 51.73346, 51.73788, 51.71745, 51.69461, 51.67548, 51.68106)
xy <- cbind(x,y)
p1 <- Polygon(xy)
poly <- SpatialPolygons(list(Polygons(list(p1), ID = "a")), proj4string=CRS("+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs"))
```

```{r, echo = FALSE}
plot(poly)
```

Generate random points with a minimum distance between each point. The rSSI function uses an inhibition distance based on a Simple Sequential Inhibition point process. I used it here for simplicity; other limited point-generation methods could be used instead.

```r
samp1 <- rSSI(0.0025, 10, poly) 
samp1 <- cbind(samp1$x, samp1$y) # Extract coordinates
samp1 <- SpatialPoints(samp1, crs(poly)) # Transform to SpatialPoints shapefile
```

```{r echo = FALSE}
plot(poly)
plot(samp1, add = TRUE)
```

Check that the minumim threshold has not been violated.

```r
spDists(samp1)*1000
```

```{r echo = FALSE}
spDists(samp1)*1000
```

Extract point coordinates and join to SpatialPoints object to create SpatialPointsDataFrame

```r
df <- data.frame(x = samp1@coords[,1], y = samp1@coords[,2]) 
samp1 <- SpatialPointsDataFrame(samp1, df)
head(samp1@data)
```
We might as well go ahead and map these to a Google Map image.

```{r echo = FALSE}
bbx <- c(left=-3.664351,bottom=51.631480,right=-3.467390,top=51.747602)
tre <- get_googlemap(center = c(lon=mean(bbx[c(1,3)]), lat=mean(bbx[c(2,4)])),
                        zoom =12, maptype = "satellite")
sDat <- data.frame(samp1@coords)
ggmap(tre) + geom_polygon(aes(x=x, y=y), data=poly, fill="red", alpha=.5) + 
  geom_point(aes(x = sDat[,1], y = sDat[,2]), data = sDat, col="orange")
```


These steps have been condensed into one function, Rpoints, below. It includes an additional step - the extraction of the lowest non-zero vaue from the distance matrix, which is then added to the dataframe.

```{r eval = FALSE}
# Function to generate random points within a polygon buffered by a minimum distance
# Attaches a dataframe of distances-between-points
#
# d = inherited distance (see rSSI documentation)
# n = number of points
# p = polygon
#
# E.g.
# spat <- Rpoints(d = 0.05, n = 50, p = poly1)
Rpoints <- function(d = 0.0025, n = 10, p){
  s <- rSSI(d, n, poly) 
  y <- SpatialPoints(cbind(s$x, s$y), crs(p))
  t <- data.frame(spDists(y)*1000)
  d <- data.frame(dist = apply(t, 1, function (v) min(v[v > 0])), # Collapse distances to non-zero minimum
                   x = y@coords[,1],
                   y = y@coords[,2])
  o <- SpatialPointsDataFrame(y,d)
}
```
